### 2017-10-05 Tex/PaperPM.tex
Additional files from Erik and Patric

%%
%%  Beam Loading Paper '17
%% ***************************
%%  Veronica K. Berglyd Olsen
%%

\documentclass[aps,prstab,reprint,amsmath,amssymb,groupedaddress,onecolumn]{revtex4-1}
\bibliographystyle{apsrev4-1}

\usepackage{hyperref}
\hypersetup{colorlinks=true, citecolor=blue, urlcolor=blue, linkcolor=blue}

\usepackage{graphicx}

% Custom Commands
%*****************

% Text
%~ \newcommand{\eq}[1]{Eq. \ref{#1}}
%~ \newcommand{\fig}[1]{Fig. \ref{#1}}
%~ \newcommand{\tbl}[1]{Table \ref{#1}}

% Math
\newcommand{\unit}[1]{\,\mathrm{#1}}
\newcommand{\funit}[2]{\,\mathrm{#1}/\mathrm{#2}}
\newcommand{\mexp}[1]{\mathrm{e}{#1}}
\newcommand{\nexp}[1]{\times 10^{#1}}

% ******************************************************************************************************************** %
\begin{document}
% ******************************************************************************************************************** %

%Patric general comments
%energy labels not consistent between figs 7, 9 and 8
%labels for the fractional charge is too complicated on Figs 7,8,9. it needs to be explained in the caption and/or in the text. "On the left axis we plot the fraction of that charge whose emittance remains .... "
%Add scale for grey color on Fig. 2
%Fig. 3 has a "v=c" and an arrow that are only on that figure ...
%It is not clear to me what the conclusions/outcome of the last section are ....

%~ Eric's Comments
%~ Sec N: Beam loading [in this section we describe the interesting physics, based on a single parameters set]
%~ - describe proton beam wake structure [similar to that of a modulated proton beam] (simulation results to show:
%~   unloaded wake)
%~ - by varying the current of the electron beam, the electron beam may load the wake and the longitudinal field can be
%~   flatten and energy spread reduced [well known result, Tzoufras] 
%~ - the electron beam blows out the remaining plasma electrons. The transverse fields are dominated by the linear
%~   focusing fields originating from the ion background, leading to emittance preservation of the part of the electron
%~   beam inside the bubble [loading of a quasi-linear wake, and still get emittance preservation: this is new]
%~   (simulation results to show: loaded wake, with bubble - many different aspects of this, trans., long. field,
%~   electron densities etc. )
%~ - the acceleration and the transverse emittance stable on long timescales (simulation results to show: transverse and
%~   longitudinal phase spaces, space long simulation)
%~ - in this regime, emittance preservation is robustness to drive-witness offset [this is new] (simulation results to
%~   show: transverse and longitudinal phase spaces, long offset simulation)
%~ Sec N+1: Parameter optimization, or, parameter scaling [in this section we describe how to optimize the electron
%~ beam, and preferably some scalings]
%~ - how to select electron beam parameters:
%~ - longitudinal parameters bunch length and current (simulations results to show: overloaded/underloaded cases)
%~ - transverse parameters of electron emittance / matching [too large, outside bubble.  Do we see increased head
%~   erosion -> faster decay]?  (simulations results to show: evolution of high emittance vs low emittance cases)
%~ (- plasma density?)

%~ Patric's Comments
%~ - Self-modulation as a driver
%~ - SM does not reach blow out
%~ - SM does not preserve emittance and produces large energy spread and parameters may evolve along the accelerator,
%~   not desirable
%~ - Beam loading for narrow energy spread
%~ - Beam loading in linear regime (old papers from SLAC in the 80's, then Katsouleas for PWFA) and nonlinear regime
%~   (Tzoufras) well know
%~ - Propose "new regime" in the SM case, beam loading and blowout to get emittance preservation
%~ - Limitation as for PWFA driver, must use some of the W bunch to reach the blowout, it is kind of like the early SLAC
%~   experiments (84GeV) where the W-bunch is like the D-bunch in these experiments, but in a plasma prepared by the p+
%~   bunch and the SM
%~ - maybe also something about the fact that 200pC in a 60µm bunch corresponds to a 1kA current, much larger than the
%~   p+ bunch current, because of the fact that the wakefields are driven my multiple bunches?
%~ As for N+2, I am a firm supporter of on-axis injection ... and most of the issue with that has to do with the plasma
%~ ramp, otherwise it would be a non-issue (assuming of course that one can split the plasma in two sections). It is in
%~ itself a topic of research, so I think we can only state a few things ...

\title{Emittance preservation of an electron beam in a loaded quasi-linear plasma wakefield}

\author{Veronica K. Berglyd Olsen}
\email[]{v.k.b.olsen@cern.ch}

\author{Erik Adli}
\affiliation{University of Oslo, Oslo, Norway}

\author{Patric Muggli}
\affiliation{Max Planck Institute for Physics, Munich, Germany}
\affiliation{CERN, Geneva, Switzerland}

\date{\today}

\begin{abstract}
We investigate beam loading and emittance preservation for a high-charge electron beam being accelerated in quasi-linear
plasma wakefield driven by a short proton beam. The structure of the studied wakefields are similar to those of a long, modulated
proton beam, such as the AWAKE proton driver.  We show that by %
properly choosing the electron beam parameters and exploiting two well known effects, beam loading of the wake field and full blow out
of plasma electrons by the accelerated beam, the electron beam can %PM be accelerated
gained large amounts of energy, have a narrow final energy spread (\%-level) and without significant emittance growth.
\end{abstract}

\maketitle

% ******************************************************************************************************************** %
\section{Introduction}\label{S:I}
% interested from SMI/AWAKE
% requirements for AWAKE Run 2
% ******************************************************************************************************************** %

Beam driven plasma wakefield accelerators \cite{chen:1985} have the potential to offer compact linear accelerators with high energy
gradients, and have been of interest for several decades. With a relativistic charged particle drive
beam travelling through a plasma, a strong wakefield is excited, that can be loaded by a trailing witness beam. When the
witness beam optimally loads the wakefield, an increase in absolute energy spread can be kept small. %to a minimum. %PM
The concept
has been demonstrated experimentally in the past using electron drive beams accelerating electron witness beams
\cite{rosenzweig:1988, blumenfeld:2007, kallos:2008} ***add [M. Litos and 23 co-authors, Nature 515, 92?95 (2014)]***.

A major challenge with plasma wakefield accelerators is, however, to accelerate a beam while keeping both energy spread and
emittance small. %growth to a minimum. PM
In the well described linear regime, valid when the beam density $n_{b}$ is much smaller than the
plasma density $n_{0}$, a non-linear transverse focusing force causes emittance growth of the witness beam. The beam
%will PM
also sees a transversely and longitudinally varying accelerating field causing a spread in energy after the beam has
been accelerated \cite{katsouleas:1987}. In the non-linear regime, where $n_{b} > n_{0}$, a bubble is formed by the
transverse oscillations of the plasma electrons, %forming PM
gathering in a sheet around an evacuated area filled with only ions. The
ions, assumed stationary, form an uniform density ion channel creating a focusing force that varies linearly with
radius. This %produces a  PM
focusing force %that PM
preserves emittance \cite{rosenzweig:1991}.
% EA: I put Rosenzweig original blow out paper as citation here instead of, \cite{lu:2006-1, lu:2006}.

In this paper we present simulation results showing how emittance preservation of a high charge density witness beam can be ensured, when accelerated by a proton drive beam producing quasi-linear wakefields \cite{rosenzweig:2010}.  By quasi-linear wakefields we mean mean wakefields with only partial blow out of the plasma electrons in the bubble.
%~ Patric's version is implemented:
The key idea is to have enough charge in the witness beam to at the same time load the wakefield to produce low
relative energy spread
%$\Delta E/E$, 
and completely blow out the electrons left in the bubble after the beam to reach conditions that preserves emittance.
The results have importance for the preparation of AWAKE Run 2 and possibly other applications in the quasi-linear regime.

% ******************************************************************************************************************** %
%\subsection[\label{S:I:SMI}]{Self-modulation as a Driver}
%~ Patric's Comments
%~ - Self-modulation as a driver
%~ - SM does not reach blow out
%~ - SM does not preserve emittance and produces large energy spread and parameters may evolve along the accelerator,
%~   not desirable
% ******************************************************************************************************************** %

% 	EA comment: is this part relevant?  Feels out of place.  We repeat this physics below when discussing proton beam SMI.  
%A train of electron drive bunches with a separation $\lambda_{pe}$ and a length $l_{b} \ll \lambda_{pe}$, where the
%plasma wavelength $\lambda_{pe} = 2\pi c/\omega_{pe}$ and the plasma frequency $\omega_{pe} = (n_{0} e^{2} / m_{e}
%\epsilon_{0})^{1/2}$, produces a field $E_{z}$ that increases for each drive bunch \cite{chen:1985}. A tailing witness
%beam loading the peak accelerating phase of this field will gain energy from the wakefield from all of the drive bunches
%when they are properly phased with respect to the wakefields. Acceleration of an electron witness beam driven by two
%electron drive beams has been demonstrated at Brookhaven National Laboratory \cite{muggli:2011}.

% ******************************************************************************************************************** %
\subsection{AWAKE Run 2}\label{S:I:AWAKE}
% ******************************************************************************************************************** %

The energy carried by electron drive bunches used in previous plasma wakefield experiments have been no more than
$100\unit{J}$ and the propagation length typically no more than $1\unit{m}$ \cite{blumenfeld:2007,caldwell:2009}.  For high-energy physics application a higher total beam energy is often desired, for instance, the energy of a high-charge electron bunch accelerated to $1\unit{TeV}$ with $1\nexp{10}$ electrons, similar to the beam that could be produced by the
international linear collider, is $1.6\unit{kJ}$. Using electron beams 
as drivers a large number of plasma stages is required to reach %PM
an energy of a $\unit{kJ}$ for the accelerated bunch. However, staging plasma accelerators without reducing the
effective gradient and spoiling the beam quality is challenging \cite{steinke:2016,lindstrom:2016}.

Proton beams available at CERN carry significantly more energy than available electron beams, $19\unit{kJ}$ for the SPS beam \cite{gschwendtner:2016},
allowing for much longer plasma wakefield accelerator stages. The SPS beam is orders of magnitude longer
than the plasma wavelengths needed for such applications, and the beam %will PM
does not drive a strong wake \cite{gschwendtner:2016}.  By letting the proton beam undergo
self-modulation before injecting the witness beam into accelerating structure, %most PM
stronger wakefields are %formed PM
excited. The self-modulation of the beam is
produced by the transverse fields generated by the beam acting upon itself, causing regions of the beam to rapidly
defocus \cite{kumar:2010}. The modulation frequency is close to that of the plasma, and produces a train of short proton
bunches along the beam axis with a surrounding halo of defocused particles. This train of bunches resonantly %produces
drives
wakefields to large %r PM 
amplitudes.

AWAKE at CERN is a proof of concept proton beam driven plasma wakefield
accelerator experiment \cite{awake_collaboration:2014}, currently in its first phase of operation. The experiment uses a $400\unit{GeV}$ proton beam delivered by the
SPS as its driver, and a single $10\unit{m}$ plasma stage with a nominal plasma density of $7\nexp{14}\unit{cm}^{-3}$
\cite{gschwendtner:2016}. This plasma density corresponds to $\lambda_{pe} = 1.26\unit{mm}$ -- matched to the transverse
size of the SPS proton beam such that $k_{pe}\sigma_{x,y} = 1$, where $k_{pe} = 2\pi/\lambda_{pe}$ is the plasma wave
number, %PM.
$\lanmda_{pe}=2\pi c/\omega_{pe}$ and $\omega_{pe}=\left(n_0e^2/m_e\epsilon_0\right)^{1/2}$ is the plasma electron angular frequency. %

The aim of the first phase of the experiment is to demonstrate self-modulation of the proton beam. The aim of
the second phase, in 2018, is to sample the wakefield with a long electron beam $\simeq\lambda_{pe}$. The study presented
here has relevance for Run 2 \cite{adli:2016}, starting in 2021 after the LHC long shut down 2, and aims to demonstrate acceleration of a short electron beam, $\ll\lambda_{pe}$,
to high energy and with a minimal increase in emittance and absolute energy spread.

\begin{figure}[hbt]
    \includegraphics[width=0.99\linewidth,trim={1mm 2mm 1mm 2mm},clip]{figures/figAWAKE}
    \caption{\label{Fig:AWAKER2} A simplified illustration of the experimental setup for AWAKE Run 2. The SPS proton
        beam undergoes self-modulation in the first plasma section. The electron witness beam is injected into the
        accelerating structure, and undergoes acceleration in the second plasma section
        \cite{berglyd_olsen:2015, adli:2016}.}
\end{figure}

The plans for AWAKE Run 2 proposes to use two plasma sections, illustrated in figure \ref{Fig:AWAKER2}. The first
section of about $4\unit{m}$ is the self-modulation stage where the proton beam undergoes self-modulation without the electron
beam present. The electron witness beam is then injected into the modulated proton beam before section two, where it
undergoes acceleration. The self-modulated proton beam does not produce a fully non-linear wakefield, and
 therefore not all plasma electrons are evacuated from the plasma bubble. The result is that the focusing force does not
increase linearly with radius %PM
and the accelerated beam emittance is not preserved.

%   EA: Moved to "witness beam"
%The small size of the accelerating structure produced by the self-modulated beam also
%puts constraints on the size of the electron witness beam, and thus its charge and current, if at the same time we wish
%to prevent large energy spread. By matching the transverse size of the beam at a given emittance to the plasma density,
%we prevent large amplitude oscillations which may cause additional energy spread as well as emittance growth.

%    EA: I kept Patric's version, but moved it to the end of the introduction
%~ Eric's version:
%~ The key idea in this paper is that while the proton beam wake is not fully blown-out, and thus leading to non-linear
%~ focusing forces, the electron beam, if intense enough, may load the field and at the same time create its own ion
%~ bubble. As will be shown, exploiting these two well known effects, the electron beam can be accelerated for long
%~ distances without significant emittance growth.

% ******************************************************************************************************************** %
\section[\label{S:M}]{Method}
%- explain analogy of short-bunch, SMI case, by referring to earlier Veronica-papers
%- simulation setup PIC/quickpic OS
% ******************************************************************************************************************** %

The focus in this study is the beam loading of the wakefields driven by the proton beam.   Studies of self-modulated proton bunches show that the proton bunch evolve as it propagates through a uniform plasma [K. V. Lotov, Phys. Plas. 18, 024501 (2011], but that small variations in the plasma profile the modulation, and thus the wake fields, may be stabilized over long distances [K. V. Lotov, Phys. Plas. 18, 024501 (2011), K. V. Lotov, Phys. Plas. 22, 103110 (2015), A. Caldwell and K. Lotov, Phys. Plas. 5, 103101 (2011)].   To study the witness beam evolution in a stable wake, independent of the dynamics of the self-modulation, %PMwe define here
here we use a single, non-evolving proton bunch as driver.  The proton bunch parameters are chosen so that key features of the wake - the plasma electron density in the wake and the longitudinal electric field - are the same as %PM
in the wake of a self-modulated proton beam with AWAKE baseline parameters \cite{gschwendtner:2016}.  Both the proton beam and the witness have Gaussian %PM
longitudinal charge distribution and bi-Gaussian transverse charge distributions.

We have previously studied the beam loading in a proton beam wake using the full PIC code Osiris \cite{fonseca:2002}
%using PM
and withs 2D cylindrical-symmetric simulations.  The studies \cite{berglyd_olsen:2015, berglyd_olsen:2016} primarily looked at beam loading, energy gain and energy spread, as well as different approaches to create a stable drive beam structure based on previous self-modulation studies.  In order to study the witness beam emittance evolution we adopt the recently released open source version of QuickPIC \cite{huang:2006, an:2013}. QuickPIC is a fully relativistic 3D quasi-static PIC code. It does not suffer from
the numerical Cherenkov effect that full PIC codes do \cite{godfrey:1974,lehe:2013}, making it a well suited tool to
study emittance preservation.  All simulation results in this paper are %PMcalculated
obtained using QuickPIC open source.

%~ In order to evaluate the quality of the accelerated bunch it was also necessary to study emittance evolution. However,
%~ full PIC codes are vulnerable to numerical growth of emittance caused by the ``numerical Cherenkov effect''
%~ \cite{godfrey:1974}. This is a know issue with the Yee EMF solver which causes the phase velocity of electromagnetic
%~ fields to be lower than $c$ while the beam moves very close to $c$. The effect can be mitigated somewhat by replacing
%~ the Yee solver with a solver designed by Lehe \cite{lehe:2013}. Our simulations showed that this numerical effect is
%~ still prominent in the high density regions of our very compact electron bunch, making it difficult to distinguish
%~ emittance growth from the physics from that originating from numerical error.

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/plasmaDenTWake}
    \caption{\label{Fig:PlasmaDenTWake} %PMThe p
    Plasma density in grey with the drive beam (blue) and the witness beam (red)
        superimposed. The line plot indicates the transverse wakefield gradient $dW_{x}/dx$ where
        $W_{x} = E_{x} - v_{b} B_{y}$, evaluated along the beam axis.
        %PM
        Beams move to the left.}
\end{figure}

% ******************************************************************************************************************** %
\subsection{Drive Beam Parameters}\label{S:M:Setup}
%~ Patric's Comments
%~ - Self-modulation as a driver
%~ - SM does not reach blow out
%~ - SM does not preserve emittance and produces large energy spread and parameters may evolve along the accelerator,
%~   not desirable
% ******************************************************************************************************************** %

%~ In these simulations we use a single proton drive beam that sets up a wakefield comparable to that which we expect to
%~ see from the self-modulated SPS beam in the second plasma cell of AWAKE Run 2 (Figure \ref{Fig:AWAKER2}).

%PM The baseline AWAKE drive beam current is insufficient to reach the non-linear regime and produce a plasma bubble. 
The modulation process used in AWAKE does not reach the fully non-linear regime and thus does not produce a bubble void of plasma electrons. When
the SPS beam, containing $3\nexp{11}$ protons \cite{gschwendtner:2016}, enters the second plasma stage (figure \ref{Fig:AWAKER2}), the peak
electric field is expected to be about $500\funit{MV}{m}$. The plasma electron density are only depleted to around $65\%$ of
nominal plasma density at the point where we inject the electron beam \cite{awake_collaboration:2016}.  The plasma electron density depletion and the peak field is replicated using a single %PM proton 
bunch with $1.46\nexp{10}$ protons (or $2.34\unit{nC}$), a %PM bunch
length $\sigma_{z} = 40\unit{\mu m}$ ($7\unit{kA}$), and a transverse size $\sigma_{x,y,proton} = 200\unit{\mu m}$.   The peak density is $0.83\cdot n_{0}$ results in a quasi-linear wake. %PM The beam energy is $400\unit{GeV}$.  
To avoid transverse evolution of the proton driver, emulating the stable propagation of the self-modulated proton beam [K. V. Lotov, Phys. Plas. 18, 024501 (2011), K. V. Lotov, Phys. Plas. 22, 103110 (2015), A. Caldwell and K. Lotov, Phys. Plas. 5, 103101 (2011)] we freeze the transverse evolution of the equivalent proton bunch by increasing the particles mass by six orders of magnitude.

% ******************************************************************************************************************** %
\subsection{Witness Beam Parameters}\label{S:M:Setup}
% ******************************************************************************************************************** %

%The small size of the accelerating structure produced by the self-modulated beam also
puts constraints on the size of the electron witness beam, and thus its charge and current, if at the same time we wish
to prevent large energy spread.  
%PM not sure what the point of the above sentence is ...
In order to prevent large amplitude oscillations %PM
of the witness beam particles, which may cause additional energy spread as well as emittance growth, we %PMstudy
consider a witness beam matched to the plasma density.  
%The beam matching relation is given by
%\begin{equation}
%    \beta = \frac{\sigma_{x,y}^{2}\gamma}{\epsilon_{N}}
%          = \frac{c}{\omega_{pe}}\sqrt{2\gamma}, \label{EQ:MatchedB}
%    % = c\sqrt{\frac{m_{e}\varepsilon_{0}}{n_{pe}e^{2}}2\gamma}
%\end{equation}
%where $\beta$ is the Twiss parameter. It then follows that the transverse profile of the beam must satisfy
%PM Not sure what the point of the above lines is ....
The matched beam transverse size is:
\begin{equation}
%    \sigma_{x,y}^{2} = c\epsilon_{N}\sqrt{2\frac{m_{e}\varepsilon_{0}}{n_{pe}e^{2}\gamma}} \label{EQ:MatchedR}
    \sigma_{x,y} =\left(c^2\epsilon_{N}^2m_{e}\varepsilon_{0}n_{pe}e^{2}\gamma}\right)^{1/4} \label{EQ:MatchedR}
\end{equation}

We %furthermore PM
assume an initial normalized emittance of $\epsilon_{N} = 2\unit{\mu m}$. This emittance is possible to produce with a standard RF-injector, while at the same time yielding a sufficient narrow beam. % EA: comment, perhaps put a ref here [***ref].

Beam loading of a short witness beam is sensitive to its position relative to the electric field \cite{tzoufras:2009} %PM
as well as to its dephasing with respect to the wakefields when at low energy.
%PM A low gamma witness beam can easily drift out of phase with the drive beam wakefield. The drift of the beams is proportional to $\gamma^{-2}$,  
To eliminate %initial 
de-phasing of the witness beam, % PM
the initial beam energy %PMwas
is set such
that $\gamma_{eb} = \gamma_{pb} = 426.3$, giving an energy of $217\unit{MeV}$.  A lower initial energy is likely to be
sufficient for AWAKE Run 2 injection. 
% EA: we have studied this to some extent, but unfortunately we do not have a good refence for the last statement, no details in \cite{berglyd_olsen:2015}.

Eq. (\ref{EQ:MatchedR}) yields a transverse size of $\sigma_{x,y}$ of $5.25\unit{\mu m}$, narrow compared to the drive beam $\sigma_{x,y,proton} = 200\unit{\mu m}$.   The bunch length was set to $\sigma_{z}=60\unit{\mu m}$, based on earlier %PM
beam loading studies \cite{berglyd_olsen:2016}.  The charge %PM was
is adjusted to $100\unit{pC}$ for optimal beam loading, as discussed in the next section. We refer to the defined drive beam and witness beam parameter set as the {base case}.

%   EA: the discussing of charge could better be discussed in the "beam loading chapter" - the information has been moved to there
% the charge required to load the field is 
% density of the electron beam reaches that of the plasma density for a bunch charge of a few $\unit{pC}$.   
% but reaches optimal beam loading at around $100-200\unit{pC}$. 
%The implication here being that the witness beam produces its own non-linear
%wakefield already at the head of the beam. The majority of the electrons within it will therefore see a linear focusing
%force preventing the bulk of the beam from undergoing emittance growth.

% ******************************************************************************************************************** %
\subsection{Simulation parameters}\label{SIM}
% ******************************************************************************************************************** %

The relatively small size of the witness beam %PM compared to the proton beam put some restrictions on the transverse resolution of the witness beam as we need a simulation box wide enough to contain the larger beam while resolving the narrower one (see figure \ref{Fig:PlasmaDenTWake}). 
puts constraints on the transverse cell size and number in the simulations: small size to resolve the small electron beam and large number of to resolve the large proton eam and the wakefields. %
We use %PMd 
a transverse grid cell size of $1.17\unit{\mu m}$, and of
$2.34\unit{\mu m}$ for the longitudinal grid cells for the simulations presented in section \ref{S:BL}. The witness beam
was simulated with $16.8\nexp{6}$ non-weighted particles.  ***NN*** plasma particles were simulated per grid cell.
%   EA: Veronica to fill in NN above

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamLoading}
    \caption{\label{Fig:BeamLoading} The top%PM plot shows the 
:unloaded longitudinal electric field with no witness beam
        (dashed blue line) and %PMthe 
        loaded field (whole blue line) along the beam axis. The beam density along the axis
        for both beams are shown in red. %PM The bottom plot compares the 
        Bottom plot: plasma densities along the beam axis for a drive
        beam with no witness beam (dashed green line), witness beam with no drive beam (dash-dotted green line), and
        both beams present (whole green line). $\xi = z - tc$ is the position in the simulation box, moving towards the
        left.}
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamPhaseSpace}
    \caption{\label{Fig:BeamPS} %PMThe p
    Phase space charge distribution of a $100\unit{pC}$, $60\unit{\mu m}$ long witness
        beam after $4\unit{m}$ of plasma. The mean momentum is $1.67\unit{GeV/c}$ with an RMS energy spread of
        $87\unit{MeV/c}$ ($5.2\%$) for the full beam.}
\end{figure}

\begin{figure*}[hbt]
    \includegraphics[width=0.495\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamEmittance}
    \includegraphics[width=0.495\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamEmittanceOffset}
    \caption{\label{Fig:BeamEmitt} Beam density in blue along the beam axis for an on-axis beam with respect to the
        drive beam axis (left), and an offset beam (right) with an offset of one $\sigma_{x} = 5.24\unit{\mu m}$ in the
        x-plane. At four different positions $z$ in the plasma stage. The red lines show a moving window calculation of
        transverse normalised emittance. The moving window calculation uses longitudinal slices of
        $l = 4\times\Delta\xi = 9.38\unit{\mu m}$ with a step of $\Delta\xi$. Only slices with more than $100$ macro
        particles have been included. The plasma density profile is included in green, and scaled up by a factor of
        $100$ to be visible. These simulations were run with an LHC energy drive beam of $7\unit{TeV}$}
\end{figure*}

\begin{figure*}[hbt]
    \includegraphics[width=\linewidth,trim={0mm 0mm 0mm 0mm},clip]{figures/beamFilamentationAll}
    \caption{\label{Fig:BeamFilament} Plots \textbf{a} to \textbf{f} show the transverse phase space of the offset
        electron beam at different plasma positions. Plot \textbf{g} shows the macro particle mean position, and plot
        \textbf{h} their RMS spread. Plots \textbf{a}, \textbf{b} and \textbf{c}, as well as the blue lines in plots
        \textbf{g} and \textbf{h} represent particles %PM
        not in the ion column (see Fig.~\ref{Fig:BeamEmitt}), with position $1.40\unit{\mu m} < \xi < 1.42\unit{\mu m}$. Plots
        \textbf{d}, \textbf{e} and \textbf{f}, and the red lines in plots \textbf{g} and \textbf{h} represent particles %PM
        in the ion column 
        with position $1.55\unit{\mu m} < \xi < 1.57\unit{\mu m}$.}
\end{figure*}

% ******************************************************************************************************************** %
\section{Beam Loading}\label{S:BL}
%~ Eric's Comments
%~ Sec N: Beam loading [in this section we describe the interesting physics, based on a single parameters set]
%~ - describe proton beam wake structure [similar to that of a modulated proton beam] (simulation results to show:
%~   unloaded wake)
%~ - by varying the current of the electron beam, the electron beam may load the wake and the longitudinal field can be
%~   flatten and energy spread reduced [well known result, Tzoufras]
%~ - the electron beam blows out the remaining plasma electrons. The transverse fields are dominated by the linear
%~   focusing fields originating from the ion background, leading to emittance preservation of the part of the electron
%~   beam inside the bubble [loading of a quasi-linear wake, and still get emittance preservation: this is new]
%~   (simulation results to show: loaded wake, with bubble - many different aspects of this, trans., long. field,
%~   electron densities etc. )
%~ - the acceleration and the transverse emittance stable on long timescales (simulation results to show: transverse and
%~   longitudinal phase spaces, space long simulation)
%~ - in this regime, emittance preservation is robustness to drive-witness offset [this is new] (simulation results to
%~   show: transverse and longitudinal phase spaces, long offset simulation)
%~
%~ Patric's Comments
%~ - Beam loading for narrow energy spread
%~ - Beam loading in linear regime (old papers from SLAC in the 80's, then Katsouleas for PWFA) and nonlinear regime
%~   (Tzoufras) well know
%~ - Propose "new regime" in the SM case, beam loading and blowout to get emittance preservation
%~ - Limitation as for PWFA driver, must use some of the W bunch to reach the blowout, it is kind of like the early SLAC
%~   experiments (84GeV) where the W-bunch is like the D-bunch in these experiments, but in a plasma prepared by the p+
%~   bunch and the SM
%~ - maybe also something about the fact that 200pC in a 60µm bunch corresponds to a 1kA current, much larger than the
%~   p+ bunch current, because of the fact that the wakefields are driven my multiple bunches?
% ******************************************************************************************************************** %

%  EA: the below has already been clearly spelled out in methods
%The single drive beam setup is designed to produce similar initial conditions for the witness beam as the self-modulated
%case. However, since the drive beam is prevented from significant transverse evolution, we are presented with an
%idealised case where the electron witness beam sees consistent wakefields throughout the plasma stage.

Figure \ref{Fig:BeamLoading} shows the results of QuickPIC simulations of the initial time step, for the base case parameters. The $E_{z}$-field
generated by the proton drive beam is seen as the blue line, shown with and without the
electron beam present. With a proton beam density $n_{pb} \simeq n_{0}$, %PMwe 
the wakefields are in the quasi-linear regime
\cite{rosenzweig:2010}. The dashed green line in the lower part of figure \ref{Fig:BeamLoading} shows that the on-axis
plasma density has a depletion of $67\%$, close to what we see in full scale reference simulations for AWAKE Run 2
\cite{awake_collaboration:2016}.

%~ The witness beam generates its own wakefield, which in the accelerating phase partially cancels out the $E_{z}$ field
%~ generated by the drive beam. For a finite length beam this causes the particles in the tail of the witness beam to be
%~ accelerated less than those at the front resulting in higher energy spread in exchange for higher energy transfer
%~ \cite{van_der_meer:1985}. A beam profile $\rho_{b}(\xi)$ that perfectly flattens the
%~ accelerating field within the witness beam will reduces energy spread to zero. The ideal shape is triangular or
%~ trapezoidal with the peak density in the direction of the beam velocity \cite{katsouleas:1987, tzoufras:2009}.

The witness beam generates its own wakefield which %PM partially cancels out 
loads the $E_{z}$ field generated by the drive beam.
With an ideally shaped electron beam charge profile it is possible to optimally load the field in such a way that the
accelerating field is constant along the beam \cite{katsouleas:1987, tzoufras:2009}.  Gaussian beams, as assumed in these studies, cannot
completely flatten the electric field in the tails of the charge distribution, and our base case beam therefore has a
tail in energy both at the front and the back of the beam, as illustrated in figure \ref{Fig:BeamPS}.  The bulk of
the beam, however, sees a relatively flat field.  
% EA: I rewrote this to the sentence after the comments
%we can acceleratoreHowever, with such a narrow beam when matched beam to the plasma density (equation
%\ref{EQ:MatchedB}), we are limited on how much total charge we can accelerate without reaching charge densities that
%will overload the field.
%  EA: shall we say the below?  The sentence asks for more detailed studies of the dependence of the loading of the transverse fields.
%The amounts of charge required to load the field depends both on the witness bunch length as well as the transverse size.

The %PM corresponding 
electron beam density is $n_{eb} \approx 35\times n_{0}$.  This means that the witness beam's own wakefield
is in the fully non-linear regime, where the space charge force is sufficient to blow-out all plasma electrons resulting
in the formation of an ion column %PM 
(Fig.~\ref{Fig:BeamLoading} bottom). This ion column, as is well known \cite{rosenzweig:1991}, provides a linear focusing
force on the part the electron beam within the column, and therefore prevents emittance growth for this %PMregion
part of the
beam.  This effect is shown for our base case in figure \ref{Fig:PlasmaDenTWake}.  The focusing field has a gradient of
$20\unit{kT/m}$ near the beam axis, corresponding to the matched field gradient.

%~ Due to the high charge density resulting from a very narrow beam, the witness beam's own wakefield reaches the full
%~ non-linear regime with $n_{eb} \approx 35\times n_{0}$. This second wakefield is driven by the beam's own front. The
%~ space charge rapidly becomes high enough to start expel plasma electrons from the beam axis and form the characteristic
%~ electrons sheet that defines the blowout regime. As a result the plasma region reaches full depletion close to the peak
%~ of the electron beam, which in turn generates strong focusing fields, with gradients of $20\unit{kT/m}$ near the beam
%~ axis (see figure \ref{Fig:PlasmaDenTWake}).

%~ The witness beam's self-generated bubble prevents the tailing end of the beam to see any significant emittance growth.
% EA: I cut the line below, we are still discussing the base case
%For the majority of the cases we studied, that maintained a stable accelerating structure, 
 %PMDue to the ion column focusing, 
 Simulation results show that about $70-80\%$ of the electron beam longitudinal slices retain their initial emittance after the propagation in the plasma.  Figure \ref{Fig:BeamEmitt} shows the slice emittance along the beam for the base case, sampled after propagating through $0$, $4$, $40$ and $100\unit{m}$ of plasma. 
Emittance growth mainly occurs in first few meters, and no significant emittance growth %PMwas
is observed after this for propagation lengths up to
$100\unit{m}$.  The head of the beam does not benefit from the full ion column focusing, but since the proton beam creates a quasi-linear wake, the emittance of the head of the beam still stabilizes after some time.  For the 100 meter simulation, the drive beam energy was increased to $7\unit{TeV}$ (LHC energy) to prevent de-phasing,
as de-phasing starts to become a significant effect for the SPS beam of $400\unit{GeV}$ after about $50\unit{m}$.

%~ Our configuration also shows some robustness to small electron beam transverse offsets as the bubble generated by the
%~ witness beam wakefield follows the beam itself.
%PM
So far we have considered a witness beam injected on the axis of the proton bunch. %
We now briefly examine the case of injection of a witness beam with an offset with respect to the axis of the proton bunch. %

Since the witness beam creates its own plasma bubble, the emittance of the part of the beam inside the
bubble is not affected by small transverse offsets of the witness beam with respect to the proton beam axis. This is an added benefit of this 
accelerating regime, and may ease transverse injection tolerances. Also in the offset case the emittance of the head of the beam still stabilizes after some time.
This is illustrated in figure \ref{Fig:BeamEmitt} for an electron beam offset of $1\sigma_{x}$. In this case there is a
larger initial emittance growth (see figure \ref{Fig:BeamFilament}a-f),
%PM
%THERE IS NO EMITTANCE SHOWN OR NOTHING TO COMPARE TO ... WHAT TO DO????
 but the emittance growth stops after the first
few meters (figure \ref{Fig:BeamEmitt}). This effect is likely to be greater for larger offsets as the beam oscillates
around the axis of the drive beam wakefield, as can bee seen %PMfrom 
in figures \ref{Fig:BeamFilament}%PMg-h.
g and h.

The transverse beam size within the bubble, where normalized emittance is preserved, follows the evolution given by
equation \ref{EQ:MatchedR}%PM where the plasma density is given by the ion density
,i .e., evolves to stay matched. The on-axis density of the electron
beam, as a result, increases as its gamma factor increases and its transverse size decreases. The effect can be seen in
figure \ref{Fig:BeamFilament} h. This has the potential to cause overloading of the field. However, for the base case
no significant overloading %PMwas
is observed.

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQuality}
    \caption{\label{Fig:BeamQ} Ratio of total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a
        function of initial beam charge (blue %PM
        symbols, lines), and relative energy spread of the accepted charge (red %PM
        symbols, lines), after
        $4\unit{m}$ of plasma and with an initial emittance $\epsilon_{N,0}=2\unit{\mu m}$. These are shown for four
        different $\sigma_{z}$ from $40\unit{\mu m}$ to $100\unit{\mu m}$. The detailed studies presented in beam
        loading section correspond to the square marked lines at $100\unit{pC}$.}
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQualityAbs}
    \caption{\label{Fig:BeamQAbs} Total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a function of
        initial beam charge (blue %PM
        symbols, lines), and final %PMmomentum
        energy (red %PM
        symbols, lines), after $4\unit{m}$ of plasma and with an initial emittance
        $\epsilon_{N,0}=2\unit{\mu m}$.}
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQualityEmittance}
    \caption{\label{Fig:BeamQEmit} Ratio of total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a
        function of initial emittance (blue %PM
        symbols, lines), and relative energy spread of the accepted charge (red %PM
        symbols, lines), after $4\unit{m}$
        of plasma.}
\end{figure}

% ******************************************************************************************************************** %
\section{Parameter Optimisation}\label{S:PO}
%~ Sec N+1: Parameter optimization, or, parameter scaling [in this section we describe how to optimize the electron
%~ beam, and preferably some scalings]
%~ - how to select electron beam parameters:
%~ - longitudinal parameters bunch length and current (simulations results to show: overloaded/underloaded cases)
%~ - transverse parameters of electron emittance / matching [too large, outside bubble.  Do we see increased head
%~   erosion -> faster decay]?  (simulations results to show: evolution of high emittance vs low emittance cases)
%~ (- plasma density?)
% ******************************************************************************************************************** %

The beam loading and blow out properties of the electron beam depend on a large number of parameters, including %PMthe
longitudinal profile, %PMthe 
transverse profile as well as %PMthe 
relative phasing of the proton and the electron beams.  We present a limited parameter study aimed at guiding beam parameter choices for AWAKE Run 2.
%~ We here study parameter optimization relevant for AWAKE Run 2, in order to understand the ideal injection parameters.
For electron beam to be externally injected in AWAKE Run 2 (see figure \ref{Fig:AWAKER2}) it is desired to maximize the energy gain, minimize the energy spread, maximize the charge to be accelerated, and minimize the emittance growth \cite{adli:2016}. In addition, the bunch length should be such that it is possible to
generate and transport the bunch using a compact electron injector \cite{adli:2016}. We
investigate the interdependence of these parameters in simulation by varying the electron bunch length, its charge,
and initial emittance. The results were quantified in terms of how much of the initial charge retained its initial
emittance within a $5\%$ margin. For these parameter scans we used a transverse grid cell size of $2.34\unit{\mu m}$,
and let the beams propagate through $4\unit{m}$ of plasma.
%~ We quantify the resulting emittance growth, energy spread and and energy gain. The desired injection parameters will
%~ depend on the application.

Figure \ref{Fig:BeamQ}  show the dependence of %PM emittance
charge and energy spread on bunch length and %PMbunch
incoming charge.  An initial beam
emittance of $2\unit{\mu m}$ was used.  %
%PM
Therefore, we define the fractional charge as that whose emittance remains smaller than $2.1\unit{\mu m}$. %PM We ran the scan with beam charges from 
The beam charge range $10\unit{pC}$ to $300\unit{pC}$ and %with length
the $\sigma_{z}$ range from $40\unit{\mu m}$ to $100\unit{\mu m}$.  As can be seen from figure \ref{Fig:BeamQ} %, red curves
(red curves), both the
$40\unit{\mu m}$ and the $60\unit{\mu m}$ beam %PMhas
have a well defined minimum energy spread with a beam charge
$\approx 50\unit{pC}$ and $\approx 100\unit{pC}$, respectively. Lower beam charges tend to under-load the electric field,
while higher beam charges tend to overload it. It is also clear that longer beams with respect to the accelerating phase
of the field, $\approx\lambda_{pe}/4$, %PMwill
does not optimally load the wake, thus producing a larger spread in energy.  The blue curves show the fraction of %PMthe sliced emittance where the
charge whose initial emittance is preserved (the slice%Md
  emittance for the base case is shown in figure \ref{Fig:BeamEmitt}).  As the witness beam charge increases the fraction of slices with preserved emittance increases, as expected from an earlier onset of the bubble formation and also increase in the bubble size \cite{lu:2006-1, lu:2006}.
%PM SHOULD WE ADD SOMETHING LIKE:
We note here that operation with 100\unit{pC}$ leads to a significantly larger charge (factor $\sim4$) with emittance preserved, at the expense of an increase of relative energy spread by a factor of two. %???

Figure \ref{Fig:BeamQAbs} shows the dependence of mean energy gain on bunch length and bunch charge %PM, in red curves
(red curves). The blue curves show the amount of charge in the longitudinal slices where the emittance has been preserved, as function of bunch length and bunch charge.

Figure \ref{Fig:BeamQEmit} shows how the growth in emittance and energy spread varies with initial electron beam
emittance.  In these simulations %PMthe focusing of the beam to the plasma remains matched for each emittance value.
we adjusted the witness beam radius to maintain the matching condition at each emittance. The smaller the initial emittance is, the better the emittance is preserved. There are two effects that lead
to emittance growth for high initial emittance beams; the transverse beam size may increase beyond the size of the bubble, and the beam
density may be reduced so much that the plasma electrons are no longer fully evacuated from the bubble. Emittance
values higher than a few micrometers lead to a significant increase in both emittance and energy spread.

The optimal working point will depend on the application and must be studied for each case, based on a trade off between bunch length, bunch charge and emittance preservation, taking into account also other parameters, like the plasma density.

% ******************************************************************************************************************** %
\section{Conclusion}\label{S:C}
% ******************************************************************************************************************** %

%We have studied a new plasma wakefield acceleration regime; acceleration of a strongly loaded electron beam in a quasi-linear proton driven wake. When optimally loaded an electron witness beam has sufficient charge to create a strong blow out forming a %n ion 
%bubble trailing the witness beam.  The focusing forces from the ions ensures emittance preservation for large parts of the %PM 
%witness beam. %PM, also when the electron beam has some offset with respect to the proton driver.
%Simulations show that the emittance of a large fraction of the electron beam is preserved even when it enters the plasma with and offset with respect to the proton bunch axis. %

%Using AWAKE parameters these result indicates that emittance preservation for an electron beam injected on-axis into a plasma stage, and properly optimized, may be
%feasible. Parameter studies indicate that up to a few $100\unit{pC}$ might be accelerated, for bunches of length
%$40-60\unit{\mu m}$. Such electron bunches may be generated by an injector based on a standard RF gun [S. Doebert, private communication].

%This study assumes a simplified case with respect to a self-modulated proton beam, where the wake is driven by a single, short proton bunch, and investigates an ideal, uniform plasma profile.  To fully investigate optimal AWAKE injection, further study should be performed including simulations of realistic plasma profiles combined with a self-modulated proton beam.  
% EA: we could also say "is underway" and cite Lotov et co?
%PM: NOT SURE WHAT PLASMA PROFILE WEA RE TALKING ABOUT ... I would suggest

We have devised a method to accelerate an electron witness bunch to high energy with a low relative energy spread while maintaining its incoming emittance in wakefields such that the accelerating structure is not void of plasma electrons. This is the case for the AWAKE experiment in which the wakefields are driven by a train of proton bunches produce by self-modulation of a long proton bunch. This method is in principle applicable to all experiments operating in the quasi-linear regime. Low relative energy spread and emittance preservation are achieved by choosing the electron beam parameters to load the wakefields and evacuate the remaining plasma electrons from the accelerating structure.

Parameter studies indicate that up to a few $100\unit{pC}$, up to 70\% of the incoming bunch charge, is accelerated, for bunch of lengths of $40-60\unit{\mu m}$. Such electron bunches may be generated by an injector based on a standard RF photo-emission gun [S. Doebert, private communication].

In order to use manageable computer time for simulations, this study assumes a simplified case with respect to a self-modulated proton beam, where the wake is driven by a single, short proton bunch producing similar wakefields. However, the wakefields driven by a train of bunches evolve with the ramp of a real plasma and when entering the plasma. Therefore, to be fully applicable to an experiment such as AWAKE, the study will have to be redone with more realistic parameters. However, using loading of the wakefields and the pure plasma ion column fields to produce an accelerated bunch with low relative energy spread and emittance remains applicable.

\section{Acknowledgements}\label{Ack}
% ******************************************************************************************************************** %

The simulations for this study have been performed using the open source version of QuickPIC released in early 2017 and owned
by UCLA.

These numerical simulations have been made possible through access to the Abel computing cluster in Oslo, Norway. Abel
is maintained by UNINETT Sigma2 AS and financed by the Research Council of Norway, the University of Bergen, the
University of Oslo, the University of Troms{\o} and the Norwegian University of Science and Technology. Project code:
nn9303k. Some of the simulations were also run on the student-maintained computing cluster ``Smaug'' at the University
of Oslo, Department of Physics.

% EA: not needed (checked with Jorge)
%The authors would also like to thank the OSIRIS Consortium for providing access to the OSIRIS framework. OSIRIS was used
%extensively for simulations leading to the work presented in this paper.
% EA: instead I suggest something like (also promised Jorge)
We gratefully acknowledge helpful discussions with Jorge Vieria from IST Lisbon.

% ******************************************************************************************************************** %
\bibliography{Bibliography}
\end{document}
% ******************************************************************************************************************** %